---
title: insee R package
output: 
  github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

<!-- <br> -->

<!-- [![CRAN status](https://www.r-pkg.org/badges/version/insee)](https://cran.r-project.org/package=insee) -->
<!-- [![CRAN checks](https://cranchecks.info/badges/worst/insee)](https://cran.r-project.org/web/checks/check_results_insee.html) -->
<!-- [![Codecov test coverage](https://codecov.io/gh/hadrilec/insee/branch/master/graph/badge.svg)](https://codecov.io/gh/hadrilec/insee?branch=master) -->
<!-- [![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing) -->
<!-- [![Downloads](https://cranlogs.r-pkg.org/badges/grand-total/insee)](https://cran.r-project.org/package=insee) -->
<!-- [![Downloads](https://cranlogs.r-pkg.org/badges/insee)](https://cran.r-project.org/package=insee) -->
<!-- [![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/) -->
<!-- [![Build Status](https://travis-ci.com/hadrilec/insee.svg?branch=master)](https://travis-ci.org/hadrilec/insee) -->
<!-- [![R build status](https://github.com/hadrilec/insee/workflows/R-CMD-check/badge.svg)](https://github.com/hadrilec/insee/actions) -->

<!-- <br> -->

## Overview

The insee package contains tools to easily download data and metadata from INSEE main database (BDM). 

Using embedded SDMX queries, get the data of more than 150 000 INSEE series.

Have a look at the detailed SDMX web service page with the following [link](https://www.insee.fr/en/information/2868055).

This package is a contribution to reproducible research and public data  transparency.

## Installation & Loading
```{r eval = FALSE}
# Get the development version from GitHub
# install.packages("devtools")
# devtools::install_github("hadrilec/insee")

# Get the CRAN version
install.packages("insee")

# examples below use tidyverse packages 
library(tidyverse)
library(insee)
```

## Examples & Tutorial

* [Tutorial](https://hadrilec.github.io/insee/articles/insee.html)
* [GDP growth rate](https://hadrilec.github.io/insee/articles/v2_gdp-vignettes.html)
* [Inflation](https://hadrilec.github.io/insee/articles/v3_inflation-vignettes.html)
* [Unemployment rate](https://hadrilec.github.io/insee/articles/v4_unem-vignettes.html)
* [Population by age](https://hadrilec.github.io/insee/articles/v5_pop-vignettes.html)
* [Population map](https://hadrilec.github.io/insee/articles/v6_pop_map-vignettes.html)

## How to avoid proxy issues ?
```{r proxy, eval=FALSE}
Sys.setenv(http_proxy = "my_proxy_server")
Sys.setenv(https_proxy = "my_proxy_server")
```


## Support
Feel free to contact me with any question about this package using this [e-mail address](mailto:hadrien.leclerc@insee.fr?subject=[r-package][insee]).
