---
title: "Tutorial"
author: 
- "Hadrien Leclerc"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    toc: true
    theme: architect
    highlight: github
    includes:
      # in_header: header.html
vignette: >
  %\VignetteIndexEntry{insee} 
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{kableExtra,magrittr,htmltools}
---

```{r setup,echo=FALSE, include=FALSE}
# setup chunk
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")),"true")
knitr::opts_chunk$set(purl = NOT_CRAN)
library(insee)
library(tidyverse)

embed_png <- function(path, dpi = NULL) {
  meta <- attr(png::readPNG(path, native = TRUE, info = TRUE), "info")
  if (!is.null(dpi)) meta$dpi <- rep(dpi, 2)
  knitr::asis_output(paste0(
    "<img src='", path, "'",
    " width=", round(meta$dim[1] / (meta$dpi[1] / 96)),
    " height=", round(meta$dim[2] / (meta$dpi[2] / 96)),
    " />"
  ))}
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(kableExtra)
library(magrittr)
library(htmltools)
library(prettydoc)
```

# Introduction

The [insee](https://CRAN.R-project.org/package=insee) package gathers tools to easily download data and metadata from [insee BDM database](https://www.insee.fr/en/information/2512785). 

It uses SDMX queries under the hood. Have a look at the detailed [SDMX webservice page](https://www.insee.fr/en/information/2868055).

The first version of the package was published on CRAN 2020-07-29.

## Proxy issues

<center><font color="red"> **Requirement for INSEE employees**</font></center>

In order for someone working behind a proxy server to be able to use [insee](https://CRAN.R-project.org/package=insee), it is necessary to modify system variables as follow.

```{r, message = FALSE, warning = FALSE, eval = FALSE}
Sys.setenv(http_proxy = "my_proxy_server")
Sys.setenv(https_proxy = "my_proxy_server")
```

## Installation

You can easily install [insee](https://CRAN.R-project.org/package=insee) with the following code :

```{r, message = FALSE, warning = FALSE, eval = FALSE}
install.packages("insee")
```

## Loading

You can easily load [insee](https://CRAN.R-project.org/package=insee) with the following code :

```{r, message = FALSE, warning = FALSE, eval = FALSE}
library(insee)
```

This vignette uses the [tidyverse](https://CRAN.R-project.org/package=tidyverse) package to exemplify the [insee](https://CRAN.R-project.org/package=insee) package functionalities with hands-on examples.

```{r, message = FALSE, warning = FALSE, eval = FALSE}
library(tidyverse)
```

# Functionalities

This section will give you an overview of what you can do with [insee](https://CRAN.R-project.org/package=insee).

Series have two identifiers the SDMX identifier and the so called idbank. Both can be used to download data.

## Datasets List

[INSEE BDM database](https://www.insee.fr/en/information/2512785) offers more than 200 Datasets. The <code>get_dataset_list()</code> function returns the datasets catalogue :

```{r, message = FALSE, warning = FALSE, eval = TRUE}
insee_dataset <- get_dataset_list() 
```

```{r echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}
rownames(insee_dataset) <- NULL

insee_dataset %>% 
  select(id, Name.en, Name.fr, url, n_series) %>% 
  slice(1:10) %>% 
  kable(row.names=NA) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## Series Key List

[INSEE BDM database](https://www.insee.fr/en/information/2512785) currently offers more than 140 000 Series. The <code>get_idbank_list</code> function returns the series catalogue.

```{r, message=FALSE,warning=FALSE,eval=TRUE}
idbank_list = get_idbank_list()
```

```{r echo=FALSE, message=FALSE, warning=FALSE,eval=TRUE}
rownames(idbank_list) <- NULL
idbank_list %>% 
          select(nomflow, idbank, cleFlow) %>%   
          group_by(nomflow) %>% 
          slice(1) %>% 
          ungroup() %>% 
          head(10) %>%
          kable(row.names=NA) %>% 
          kable_styling(bootstrap_options = c("striped", "hover", "condensed"))


```

## Find a series key

The best way to download data is to find the right series key (idbank), but how ? Indeed, in some cases it is not easy to understand what are the differences among series, especially for non-French speakers.
To make the search easier, the [insee](https://CRAN.R-project.org/package=insee) package provides the function <code>get_insee_title</code> to get the title related to an idbank, either in English or in French. It is not advised to use the <code>get_insee_title</code> function on the whole idbank dataset, as each SDMX query has 400-idbank limit. Then, <code>get_insee_title</code> function splits the input into several lists of 400 idbanks each and is consequently quite slow. Thus, the user should filter the idbank dataset before using the <code>get_insee_title</code> function to avoid as much as possible this bottleneck as the following example shows. 

```{r message=FALSE, warning=FALSE,eval=FALSE}
idbank_list = get_idbank_list()

idbank_list_selected =
  idbank_list %>%
  filter(nomflow == "IPI-2015") %>% #industrial production index dataset
  filter(dim1 == "M") %>% #monthly
  filter(dim5 == "INDICE") %>% #index
  filter(dim8 == "CVS-CJO") %>% #Working day and seasonally adjusted SA-WDA
  #automotive industry and overall industrial production
  filter(str_detect(dim4,"^29$|A10-BE")) %>% 
  mutate(title = get_insee_title(idbank))

idbank_list_selected

```

## Download data

### Download using idbank

The <code>get_insee_idbank</code> function should handle up to 400 idbanks but not more.
It is then advised to narrow down the idbanks list used as argument of the function.

```{r message=FALSE, warning=FALSE,eval=FALSE}
library(tidyverse)
library(insee)

# the user can make a manual list of idbanks to get the data 
# example 1

data = get_insee_idbank("001558315", "010540726")

# using a list of idbanks extracted from the insee idbank dataset
# example 2 : household's confidence survey

idbank_dataset = get_idbank_list()

df_idbank = idbank_dataset %>%
  filter(nomflow == "ENQ-CONJ-MENAGES") %>%  #monthly households' confidence survey
  mutate(title = get_insee_title(idbank)) %>%
  separate(title, sep = " - ", into = paste0("title", 1:4),
           fill = "right", remove = FALSE) %>%
  filter(dim7 == "CVS") #seasonally adjusted

# dim2: IND_SYNT_CONF - Summary indicator of households' confidence

list_idbank = df_idbank %>% pull(idbank)

data = get_insee_idbank(list_idbank)

```

### Download using dataset name

For some datasets as IPC-2015 (inflation), the filter is necessary.

```{r message=FALSE, warning=FALSE,eval=FALSE}

insee_dataset <- get_dataset_list() 

# example 1 : full dataset
data = get_insee_dataset("CLIMAT-AFFAIRES")

# example 2 : filtered dataset 
# the user can filter the data
data = get_insee_dataset("IPC-2015", filter = "M+A.........CVS.", startPeriod = "2015-03")

# in the filter, the + is used to select several values in one dimension, like an "and" statement
# the void means "all" values available

# example 3 : only one series
# by filtering with the full SDMX series key, the user will get only one series
data = 
  get_insee_dataset("CNA-2014-CPEB",
                    filter = "A.CNA_CPEB.A38-CB.VAL.D39.VALEUR_ABSOLUE.FE.EUROS_COURANTS.BRUT",
                    lastNObservations = 10)

```

# Examples 

* [GDP growth rate](https://hadrilec.github.io/insee/articles/2_gdp-vignettes.html)
* [Inflation](https://hadrilec.github.io/insee/articles/3_inflation-vignettes.html)
* [Unemployment rate](https://hadrilec.github.io/insee/articles/4_unem-vignettes.html)
* [Population by age](https://hadrilec.github.io/insee/articles/5_pop-vignettes.html)
* [Population map](https://hadrilec.github.io/insee/articles/6_pop_map-vignettes.html)


# Support
Feel free to contact me with any question about this package using this [e-mail address](mailto:leclerc.hadrien@gmail.com?subject=[r-package][insee]).

# Disclaimer

This package is in no way officially related to or endorsed by INSEE.

