---
title: "package R : insee"
author: 
- "Hadrien Leclerc"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    toc: true
    theme: architect
    highlight: github
    includes:
      # in_header: header.html
vignette: >
  %\VignetteIndexEntry{insee} 
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{kableExtra,magrittr,htmltools}
---

```{r setup,echo=FALSE, include=FALSE}
# setup chunk
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")),"true")
knitr::opts_chunk$set(purl = NOT_CRAN)
library(insee)
library(tidyverse)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(kableExtra)
library(magrittr)
library(htmltools)
library(prettydoc)
```

# Introduction

The [insee](https://CRAN.R-project.org/package=insee) package gathers tools to easily download data and metadata from [insee BDM database](https://www.insee.fr/en/information/2512785). 

It uses SDMX queries under the hood. Have a look at the detailed [SDMX webservice page](https://www.insee.fr/en/information/2868055).

The first version of the package was published on CRAN 2020-07-29.

## Proxy issues

<center><font color="red"> **Requirement for INSEE employees**</font></center>

In order for someone working behind a proxy server to be able to use [insee](https://CRAN.R-project.org/package=insee), it is necessary to modify system variables as follow.

```{r, message=FALSE,warning=FALSE,eval=FALSE}
Sys.setenv(http_proxy = "my_proxy_server")
Sys.setenv(https_proxy = "my_proxy_server")
```

## Installation

You can easily install [insee](https://CRAN.R-project.org/package=insee) with the following code :

```{r, message=FALSE,warning=FALSE,eval=FALSE}
install.packages("insee")
```

## Loading

You can easily load [insee](https://CRAN.R-project.org/package=insee) with the following code :

```{r, message=FALSE,warning=FALSE,eval=FALSE}
library(insee)
```

This vignette uses the [tidyverse](https://CRAN.R-project.org/package=tidyverse) package to exemplify the [insee](https://CRAN.R-project.org/package=insee) package functionalities with hands-on examples.

```{r, message=FALSE,warning=FALSE,eval=FALSE}
library(tidyverse)
```

# Functionalities

This section will give you an overview of what you can do with [insee](https://CRAN.R-project.org/package=insee).

Series have two identifiers the SDMX identifier and the so called idbank. Both can be used to download data.

## Datasets List

[INSEE BDM database](https://www.insee.fr/en/information/2512785) offers more than 200 Datasets. The <code>get_dataset_list()</code> function returns the datasets catalogue :

```{r, message=FALSE,warning=FALSE,eval=NOT_CRAN}
insee_dataset <- get_dataset_list() 
```

```{r echo=FALSE, message=FALSE, warning=FALSE,eval=NOT_CRAN}
rownames(insee_dataset) <- NULL
insee_dataset %>% 
  select(id, Name.en, Name.fr, url, n_series) %>% 
  slice(1:10) %>% 
  kable(row.names=NA) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## Series Keys List

[INSEE BDM database](https://www.insee.fr/en/information/2512785) currently offers more than 140 000 Series. The <code>get_idbank_list</code> function returns the series catalogue.

```{r, message=FALSE,warning=FALSE,eval=NOT_CRAN}
idbank_list = get_idbank_list()
```

```{r echo=FALSE, message=FALSE, warning=FALSE,eval=NOT_CRAN}
rownames(idbank_list) <- NULL
idbank_list %>% 
          select(nomflow, idbank, cleFlow) %>%   
          group_by(nomflow) %>% 
          slice(1) %>% 
          ungroup() %>% 
          head(10) %>%
          kable(row.names=NA) %>% 
          kable_styling(bootstrap_options = c("striped", "hover", "condensed"))


```

## Find a series key

The best way to download data is to find the right series key (idbank), but how ? Indeed, in some cases it is not easy to understand what are the differences among series, especially for non-French speakers.
To make the search easier, the [insee](https://CRAN.R-project.org/package=insee) package provides the function <code>get_insee_title</code> to get the title related to an idbank, either in English or in French. It is not advised to use the <code>get_insee_title</code> function on the whole idbank dataset, as each SDMX query has 400-idbank limit. Then, <code>get_insee_title</code> function splits the input into several lists of 400 idbanks each and is consequently quite slow. Thus, the user should filter the idbank dataset before using the <code>get_insee_title</code> function to avoid as much as possible this bottleneck as the following example shows. 

```{r message=FALSE, warning=FALSE,eval=NOT_CRAN}
idbank_list = get_idbank_list()

idbank_list_selected =
  idbank_list %>%
  filter(nomflow == "IPI-2015") %>% #industrial production index dataset
  filter(dim1 == "M") %>% #monthly
  filter(dim5 == "INDICE") %>% #index
  filter(dim8  == "CVS-CJO") %>% #Working day and seasonally adjusted SA-WDA
  filter(str_detect(dim4,"^29$|A10-BE")) %>% #automotive industry and overall industrial production
  mutate(title = get_insee_title(idbank))

idbank_list_selected

```

## Download data

### Download using idbank

```{r message=FALSE, warning=FALSE,eval=NOT_CRAN}
library(tidyverse)

data = get_insee_idbank("001558315", "010540726")

```

### Download using dataset name

```{r message=FALSE, warning=FALSE,eval=NOT_CRAN}
library(tidyverse)

data = get_insee_idbank("001558315", "010540726")

```

# Useful examples

## GDP growth rate 

```{r message=FALSE, warning=FALSE,eval=NOT_CRAN}
library(tidyverse)

idbank_list = get_idbank_list()

df_idbank_list_selected =
  idbank_list %>%
  filter(nomflow == "CNT-2014-PIB-EQB-RF") %>% # Gross domestic product balance
  filter(dim1 == "T") %>% #quarter
  filter(dim4 == "PIB") %>% #GDP
  filter(dim6 == "TAUX") %>% #rate
  filter(dim10 == "CVS-CJO") #SA-WDA, seasonally adjusted, working day adjusted

idbank = df_idbank_list_selected %>% pull(idbank)

data = get_insee_idbank(idbank)

#plot
ggplot(data, aes(x = DATE, y = OBS_VALUE)) +
  geom_col() +
  ggtitle("French GDP growth rate, quarter-on-quarter, sa-wda") +
  labs(subtitle = sprintf("Last updated : %s", data$TIME_PERIOD[1]))

```

## Inflation

```{r message=FALSE, warning=FALSE,eval=NOT_CRAN}
library(tidyverse)
library(lubridate)

idbank_list = get_idbank_list()

df_idbank_list_selected =
  idbank_list %>%
  filter(nomflow == "IPC-2015") %>% #Inflation dataset
  filter(dim1 == "M") %>% # monthly
  filter(str_detect(dim4, "^[0-9]{2}$")) %>% # coicop aggregation level
  filter(dim6 == "INDICE") %>% # index
  filter(dim7 == "ENSEMBLE") %>% # all kinds of household
  filter(dim8 == "FE") %>% # all France including overseas departements
  mutate(title = get_insee_title(idbank))

list_idbank = df_idbank_list_selected %>% pull(idbank)

data = get_insee_idbank(list_idbank, startPeriod = "2010-01")

n_sep = str_count(data$TITLE_FR[1], " - ") + 1

data_plot = data %>%
  separate(TITLE_EN, into = paste0("title", 1:n_sep), sep = " - ", remove = FALSE, fill = "right") %>%
  mutate(title6 = case_when(is.na(title6) ~ title5,
                            TRUE ~ as.character(title6))) %>%
  mutate(title6 = substr(title6, 1 , 22)) %>%
  mutate(month = month(DATE)) %>%
  arrange(DATE) %>%
  group_by(title6, month) %>%
  mutate(growth = 100 * (OBS_VALUE / dplyr::lag(OBS_VALUE) - 1))

ggplot(data_plot, aes(x = DATE, y = growth)) +
  geom_col() +
  facet_wrap(~title6, scales = "free") +
  ggtitle("French inflation, by product categories, year-on-year") +
  labs(subtitle = sprintf("Last updated : %s", data_plot$TIME_PERIOD[nrow(data_plot)])) 

```

## Unemployment rate

```{r message=FALSE, warning=FALSE,eval=NOT_CRAN}
library(tidyverse)

dataset_list = get_dataset_list()

idbank_list = get_idbank_list()

df_idbank_list_selected =
  idbank_list %>%
  filter(nomflow == "CHOMAGE-TRIM-NATIONAL") %>% #Unemployment dataset
  mutate(title = get_insee_title(idbank)) %>%
  filter(dim2 == "CTTXC") %>% #unemployment rate based on ILO standards
  filter(dim4 == "FE") %>%  # all France including overseas departements
  filter(dim5 == 0) # men and women

list_idbank = df_idbank_list_selected %>% pull(idbank)

data = get_insee_idbank(list_idbank, startPeriod = "2000-01")

n_sep = str_count(data$TITLE_FR[1], " - ") + 1

data_plot = data %>%
  separate(TITLE_EN, into = paste0("title", 1:n_sep), sep = " - ", remove = FALSE, fill = "right")

ggplot(data_plot, aes(x = DATE, y = OBS_VALUE, colour = title2)) +
  geom_line() +
  geom_point() +
  ggtitle("French unemployment rate, by age") +
  labs(subtitle = sprintf("Last updated : %s", data_plot$TIME_PERIOD[1]))

```

## Population

```{r message=FALSE, warning=FALSE,eval=NOT_CRAN}
library(tidyverse)

dataset_list = get_dataset_list()

idbank_list = get_idbank_list()

df_idbank_list_selected =
  idbank_list %>%
  filter(nomflow == "POPULATION-STRUCTURE") %>% #population dataset
  mutate(title = get_insee_title(idbank)) %>%
  filter(dim2 == "POPULATION_1ER_JANVIER") %>% #population at the beginning of the year
  filter(dim5 == "FE") %>%  # all France including overseas departements
  filter(dim6 == 0) %>%  # men and women
  filter(dim7 %in% c("00-19", "20-59", "60-")) #age ranges

list_idbank = df_idbank_list_selected %>% pull(idbank)

data = get_insee_idbank(list_idbank)

n_sep = str_count(data$TITLE_FR[1], " - ") + 1

data_plot = data %>%
  separate(TITLE_EN, into = paste0("title", 1:n_sep), sep = " - ", remove = FALSE, fill = "right") %>%
  mutate(OBS_VALUE = OBS_VALUE / 10^6)

ggplot(data_plot, aes(x = DATE, y = OBS_VALUE, fill = title3)) +
  geom_area() +
  ggtitle("French population in millions, by age") +
  labs(subtitle = sprintf("Last updated : %s", data_plot$TIME_PERIOD[1]))

```

# Support
Feel free to contact me with any question about this package using this [e-mail address](mailto:leclerc.hadrien@gmail.com?subject=[r-package][insee]).
